Pillar 1 - Sequential Pattern Mining

### 2019-02-01 Pillar 1 - Sequential Pattern Mining.Rmd
update functions to make transactions and sequences

n_students <- d_transactions %>%
  select(sequenceID) %>%
  n_distinct()

d_support_take <- d_transactions %>%
  distinct(
    item,
    sequenceID
    ) %>%
  group_by(
    item
    ) %>%
  summarise(
    count     = n(),
    rate.take = count / n_students
    )

d_support <- d_transactions %>%
  mutate(
    grade_round = round(Grade)
  ) %>%
  group_by(
    item
    ) %>%
  summarise(
    rate.fail = mean(PF == "fail"),
    rate.low  = mean(HL == "low"),
    rate.0    = mean(grade_round <= 0),
    rate.1    = mean(grade_round <= 1),
    rate.2    = mean(grade_round <= 2),
    rate.3    = mean(grade_round <= 3),
    rate.4    = mean(grade_round <= 4),
    rate.5    = mean(grade_round <= 5),
    rate.6    = mean(grade_round <= 6),
    rate.7    = mean(grade_round <= 7),
    rate.8    = mean(grade_round <= 8),
    rate.9    = mean(grade_round <= 9),
    rate.10   = mean(grade_round <= 10)
    )

@@

  data <- data %>%
    group_by(
      sequenceID
    ) %>%
    summarise(
      list_item = list(!!enquo(item))
    ) %>%
    ungroup

  transactions <- as(
    data$list_item,
    "transactions"
    )

  return(transactions)

@@

  data <- data %>%
    filter(!is.na(eventID))%>% #this is here casue we have courses that were never taken.
    group_by(
      sequenceID,
      eventID
    ) %>%
    summarise(
      list_item = list(!!enquo(item))
    ) %>%
    arrange(
      sequenceID,
      eventID
    ) %>%
    ungroup

  sequences <- as(
    data$list_item,
    "transactions"
    )

@@

#
# Checking sequences
#inspect(head(sequences, 10))

@@

## Mining Rules
### Apriori Algorithm
```{r function clean AR}

# Transform AR from apriori into a readable data frame
clean_AR <- function(AR){

  AR %>%
    as("data.frame") %>%
    filter(
      count >= 1
    ) %>%

    # Clean variable `rules`
    mutate(
      rules = str_remove_all(rules, pattern = "[{]"),
      rules = str_remove_all(rules, pattern = "[}]")
    ) %>%
    separate(rules, into = c("lhs", "rhs"), sep=" => ") %>%

@@

We apply Apriori algorithm:
```{r my_apriori function}
# Run the apriori algorithm with desired parameters
my_apriori <- function(data, appearance = NULL){

  data %>%
    apriori(
      parameter = list(
        supp = 0,    # min support
        smax = 1,    # max support
        conf = 0,    # min confidence
        minlen = 2,  # min length of 
        maxlen = 2,  # max length of rule
        ext = TRUE,
        arem = "diff",
        aval = TRUE,
        minval = 0
        ),
      appearance = appearance,
      control = list(
        verbose = FALSE
        )
      ) %>%
    clean_AR

@@

@@

##high/low filter
SR_HL      <- my_SR(data = sequences_HL)%>% 
  filter(
    str_detect(rhs, "low"),
    str_detect(lhs, "low")
    )
```

>>>>

# Number of students
n_students <- d_transactions %>%
  select(sequenceID) %>%
  n_distinct()

# Probability of taken a course
d_support_take <- d_transactions %>%
  distinct(
    item,
    sequenceID
    ) %>%
  group_by(
    item
    ) %>%
  summarise(
    count     = n(),
    rate.take = count / n_students
    )

# Probability of failing, having a low grade of a grade less than x in a course
d_support <- d_transactions %>%
  mutate(
    grade_round = round(Grade)
  ) %>%
  group_by(
    item
    ) %>%
  summarise(
    rate.fail = mean(PF == "fail"),
    rate.low  = mean(HL == "low"),
    rate.0    = mean(grade_round <= 0),
    rate.1    = mean(grade_round <= 1),
    rate.2    = mean(grade_round <= 2),
    rate.3    = mean(grade_round <= 3),
    rate.4    = mean(grade_round <= 4),
    rate.5    = mean(grade_round <= 5),
    rate.6    = mean(grade_round <= 6),
    rate.7    = mean(grade_round <= 7),
    rate.8    = mean(grade_round <= 8),
    rate.9    = mean(grade_round <= 9),
    rate.10   = mean(grade_round <= 10)
    )

@@

  data %>%
    group_by(
      sequenceID
    ) %>%
    summarise(
      list_item = list(!!enquo(item))
    ) %>%
    ungroup %>%
    pull(
      list_item
      ) %>%
    as("transactions")

@@

  sequences <- data %>%
    filter(!is.na(eventID))%>% #this is here casue we have courses that were never taken.
    group_by(
      sequenceID,
      eventID
    ) %>%
    summarise(
      list_item = list(!!enquo(item))
    ) %>%
    arrange(
      sequenceID,
      eventID
    ) %>%
    ungroup %>%
    pull(
      list_item
      ) %>%
    as("transactions")

@@

#
# Checking sequences
# inspect(head(sequences, 10))

@@

## Mining Rules
### Apriori Algorithm

The function `my_apriori()` applies the apriori algorithm on a set of transactions with the parameters that we have chosen.
```{r my_apriori function}

my_apriori <- function(data, appearance = NULL){

  data %>%
    apriori(
      parameter = list(

        # Chosen parameters
        supp = 0,    # min support
        smax = 1,    # max support
        conf = 0,    # min confidence
        minlen = 2,  # min length of 
        maxlen = 2,  # max length of rule
        ext = TRUE,
        arem = "diff",
        aval = TRUE,
        minval = 0
        ),

      # impose restriction on what could be on the lhs and rhs of the rule
      appearance = appearance

      )

The function `clean_AR()` transforms the rules generated by the function `my_apriori()` into a readable dataframe
```{r function clean AR}

clean_AR <- function(AR){

  AR %>%
    as("data.frame") %>%

    # Exclude rules with support of `0`
    filter(
      count >= 1
    ) %>%

    # Clean variable `rules`
    mutate(
      rules = str_remove_all(rules, pattern = "[{]"),
      rules = str_remove_all(rules, pattern = "[}]")
    ) %>%
    separate(rules, into = c("lhs", "rhs"), sep = " => ") %>%

@@

The function `compute_support()` computes the support of the lhs and rhs of the rules.

```{r}
compute_support <- function(AR, type_rule){

  type_rule_enquo <- enquo(type_rule)

  AR %>%
    left_join(
      select(d_support, item, !!type_rule_enquo),
      by = c("rhs" = "item")
      ) %>%
    mutate(
      rhs.support = round(!!type_rule_enquo, 5)
      ) %>%
    select(
      - !!type_rule_enquo
      ) %>%
    left_join(
      select(d_support, item, !!type_rule_enquo),
      by = c("lhs" = "item")
      ) %>%
    mutate(
      lhs.support = round(!!type_rule_enquo, 5)
      ) %>%
    select(
      - !!type_rule_enquo
      ) %>%
    select(
      lhs, rhs, support, count, confidence, rhs.support, lift
    )

We encapsulate the three functions `my_apriori()`, `clean_AR()` and `compute_support` into the function `make_AR()`.

```{r}
make_AR <- function(data, appearance, type_rule){

  data %>%
    my_apriori(appearance = appearance) %>%
    clean_AR() %>%
    compute_support(type_rule = type_rule)

@@

AR_take <- make_AR(transactions)

@@

##high/low filter
SR_HL      <- my_SR(data = sequences_HL) %>% 
  filter(
    str_detect(rhs, "low"),
    str_detect(lhs, "low")
    )
```
